Practical Optional Types for Clojure

### 2015-02-23 overview.tex
final example

@@

\inputminted[firstline=4]{clojure}{code/demo/src/demo/eg2.clj}
\inputminted[firstline=4]{clojure}{code/demo/src/demo/eg3.clj}
\inputminted[firstline=4]{clojure}{code/demo/src/demo/eg4.clj}
\inputminted[firstline=4,lastline=15]{clojure}{code/demo/src/demo/eg5.clj}
\inputminted[firstline=4,lastline=15]{clojure}{code/demo/src/demo/eg6.clj}
\inputminted[firstline=6,lastline=23]{clojure}{code/demo/src/demo/eg7.clj}

>>>>

A multimethod in Clojure is a function with a dispatch mechanism to
of internal methods. Multimethods are created with {\clj{defmulti}}
and methods are installed via {\clj{defmethod}}.
They are called like regular functions.

\begin{minted}{clojure}
(ann rep [Any -> String])
(defmulti rep class)
(defmethod rep Number [x] (str (inc x)))
(defmethod rep Keyword [x] (str (name x)))

(rep :a) ;=> "a"
(rep 1)  ;=> "2"
\end{minted}

The \clj{class} function returns the class of an object, or \clj{nil}
if given \clj{nil}. This multimethod's dispatch function returns the class of its
argument. The underlying dispatch strategy features \clj{isa?}, a function
that is a subclassing check when provided with classes

\begin{minted}{clojure}
  (isa? Keyword Object) ;=> true
  (isa? Keyword Number) ;=> false
\end{minted}

otherwise an equality check.

\begin{minted}{clojure}
  (isa? :a :a) ;=> true
  (isa? :a 1) ;=> false
\end{minted}

So in the previous example, \clj{(rep :a)} dispatched to the method
where \clj{(isa? (class :a) v)} returned true, where \clj{v} is the dispatch value.
Since \clj{(isa? (class :a) Keyword)} is true, the second method is chosen.

The flexibility of \clj{isa?} is key to the generality of multimethods. 
For instance, we can dispatch on the value of
a map entry, even if it is not a class.

\inputminted[firstline=4,lastline=23]{clojure}{code/demo/src/demo/eg5.clj}

\clj{isa?} is further special-case with vectors: with two vectors of the
same length, it returns the logical conjunction of the pairwise applications
of \clj{isa?}.

\begin{minted}{clojure}
  (isa? [Keyword :a] [Object :a]) ;=> true
  (isa? [Keyword Keyword] [Object Object]) ;=> true
\end{minted}

We can use this to dispatch on more than one thing. Here we perform
double dispatch on the class of each argument.

\inputminted[firstline=6,lastline=23]{clojure}{code/demo/src/demo/eg7.clj}

No extra annotations are needed to follow the type-directed control flow.
For instance, the middle method is has inferred the left argument is a number
and thus safe to increment, and the right argument is a keyword and safe to
extract its name.

A dispatch value can also be the keyword \clj{:default} to specify a default
method if no preferred method is found. The static type of both arguments
remains unchanged.

@@

The final example combines everything we will cover for the rest of the paper:
multimethod dispatch, reflection resolution via type hints, Java method
and constructor calls, conditional and exceptional flow reasoning
and use of HMaps.

\inputminted[firstline=6,lastline=20]{clojure}{code/demo/src/demo/eg8.clj}
